---
title: "Evanston Open Data Research"
output:
  html_document:
    df_print: paged
---
This file is for exploratory research for Northwestern University's City of Evanston Open Data Contest: https://sites.northwestern.edu/lovedataweek/contest/

Deadline: Monday, April 1 @ 3:00 PM

```{r}
## Libraries
install.packages("ggmap")
install.packages("leaflet")
install.packages("leaflet.extras")

library(ggplot2)
library(ggmap)
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(htmlwidgets)

theme_set(theme_linedraw())
```


## Traffic Stops 2017-2018
```{r}
traffic_data <- read.csv("https://data.cityofevanston.org/resource/bbbk-a8vu.csv?$limit=35000")
levels(data$day_of_the_week) <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
```
#### Stops by Day of Week/Race
```{r}
traffic_data %>%
  ggplot() +
  geom_bar(aes(day_of_the_week)) +
  facet_wrap(~race)
```
#### Stops by Race/Beat
```{r}
traffic_data %>%
  ggplot() +
  geom_bar(aes(race)) +
  facet_wrap(~police_beat)
```
## Evanston Arrests
```{r}
arrest_data <- read.csv("https://data.cityofevanston.org/resource/jwmg-4gbx.csv")

arrest_data$loc <- as.character(arrest_data$location)

arrest_data <- arrest_data %>%
  drop_na(loc)

arrest_data$loc <- arrest_data$loc %>%
  substring(8)

arrest_data$loc <- arrest_data$loc %>%
  str_sub(1, str_length(arrest_data$loc)-1)

arrest_data_lonlat <- arrest_data %>%
  separate(col = loc, c("lon", "lat"), sep = " ")

arrest_data_lonlat$lon <- as.numeric(arrest_data_lonlat$lon)
arrest_data_lonlat$lat <- as.numeric(arrest_data_lonlat$lat)

arrest_data_lonlat <- arrest_data_lonlat %>%
  drop_na(lat)

```

## Mapping Arrests in Evanston
This is a map of arrests made in Evanston.
```{r}
leaflet() %>%
  addCircleMarkers(lng=arrest_data_lonlat$lon, lat=arrest_data_lonlat$lat, popup = arrest_data_lonlat$arrest_type) %>%
  addProviderTiles(providers$CartoDB.Positron)
```

This is a heatmap of arrests made in Evanston.
```{r}
leaflet(arrest_data_lonlat) %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addHeatmap(lng = ~lon, lat = ~lat,
             blur = 30, max = 0.05, radius = 15)
```

##All Evanston Police Activity
```{r}
pdata <- read.csv("https://data.cityofevanston.org/resource/t5g5-xgby.csv?$limit=200000")

pdata$location <- as.character(pdata$location)

pdata <- pdata %>%
  filter(location != "")

pdata$loc <- pdata$location

pdata$loc <- pdata$loc %>%
  substring(8)

pdata$loc <- pdata$loc %>%
  str_sub(1, str_length(pdata$loc)-1)

pdata <- pdata %>%
  separate(col = loc, c("lon", "lat"), sep = " ")

pdata$lon <- as.numeric(pdata$lon)
pdata$lat <- as.numeric(pdata$lat)
```
This is a heatmap of all police activity in Evanston.
```{r}
pdata %>%
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addHeatmap(lng = ~lon, lat = ~lat,
             blur = 20, radius = 8) 
```

#Tree Information
```{r}
## Read in Data
tdata <- read.csv("Trees_data.csv")

##Set Geo
tdata$location <- as.character(tdata$the_geom)

tdata <- tdata %>%
  filter(location != "")

tdata$loc <- tdata$location

tdata$loc <- tdata$loc %>%
  substring(8)

tdata$loc <- tdata$loc %>%
  str_sub(1, str_length(x$loc)-1)

tdata <- tdata %>%
  separate(col = loc, c("lon", "lat"), sep = " ")

tdata$lon <- as.numeric(tdata$lon)
tdata$lat <- as.numeric(tdata$lat)
```

## Types of Trees
```{r}
tdata %>%
  group_by(Species) %>%
  count(sort=TRUE) %>%
  head(30)

```

## Stop and Frisk

Data from https://data.cityofevanston.org/browse?q=Stop%20and%20Frisk&sortBy=relevance

```{r}
sfdata <- read.csv("Stop_and_Frisk.csv")
```

By Race
```{r}
sfdata %>%
  group_by(Race) %>%
  ggplot(aes(Race)) +
  geom_bar() +
  coord_flip() +
  scale_y_continuous(expand = c(0,0), limits = c(0, 1300)) 
  
```

By Reason
```{r}
sfdata %>%
  mutate(Reason.Code.Description = fct_rev(fct_infreq(Reason.Code.Description))) %>%
  ggplot(aes(Reason.Code.Description)) +
  geom_bar() +
  coord_flip() +
  scale_y_continuous(expand = c(0,0))
```

```{r}
sfdata %>%
  group_by(Race) %>%
  ggplot(aes(Race)) +
  geom_bar() +
  coord_flip() +
  facet_grid(~Enforcement) +
  scale_y_continuous(expand = c(0,0))
```

```{r}
sfdata %>%
  filter(grepl('STREET', Reason.Code.Description)) %>%
  ggplot(aes(Race, Age)) +
  geom_violin()
```

```{r}
sfdata %>%
  group_by(Reason.Code.Description, Enforcement) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  filter(Enforcement == "ARREST") %>%
  print(n=15)
```

```{r}
sfdata %>%
  group_by(Race, Enforcement) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  filter(Enforcement == "ARREST")
```

## Evanston Population Overall

```{r}
pop <- read.csv("https://data.cityofevanston.org/resource/j228-4hwk.csv")

```

